Queue Aware Precoder Design for Space Frequency Resource Allocation

### 2014-09-09 Report/wsrm_formulation.tex
revision - 1

In this approach, we present an iterative algorithm to solve \eqref{eqn-3} locally based on the idea of alternating optimization and successive convex approximation. Using \eqref{eq:SINR}, we can reformulate the problem defined in \eqref{eqn-3} as
\begin{IEEEeqnarray}{RCL}\label{eqn-6}
\underset{\substack{\gamma_{l,k,n},\mvec{m}{l,k,n},\\\beta_{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} & \quad & \|  \tilde{\mbf{v}}  \|_q \label{eqn-obj} \IEEEyessubnumber \\
\text{subject to}& \quad &\gamma_{l,k,n} \leq \dfrac{\left | \mvec{w}{l,k,n}^\herm \mvec{H}{l,k,n} \mvec{m}{l,k,n} \right |^2}{\beta_{l,k,n}} \triangleq f(\tilde{\mbf{u}}_{l,k,n}) \IEEEyessubnumber \label{eqn-6.2} \\
  & \quad & \beta_{l,k,n} \geq  N_0 \|\mvec{w}{l,k,n}\|^2 + \hspace{-0.75em} \sum_{(j,i) \neq (l,k)} \hspace{-0.75em} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \label{eqn-6.3} \\
  & \quad &\eqref{eqn-4.3} \IEEEyessubnumber
\end{IEEEeqnarray}
where \me{\tilde{\mbf{u}}_{l,k,n} \triangleq \{\mvec{w}{l,k,n}^\herm, \mvec{H}{b_k,k,n}, \mvec{m}{l,k,n},\beta_{l,k,n}\}} be the vector which needs to be identified for the optimal allocation. In this formulation, we relaxed the equality constraint in \eqref{eq:SINR} by the inequalities in \eqref{eqn-6.2} and \eqref{eqn-6.3}. However, this step is without loss of optimality leads to the same solution, since the inequalities in \eqref{eqn-6.2} and \eqref{eqn-6.3} are active for an optimal solution, following the same arguments as those in \cite{tran2012fast}. Intuitively, \eqref{eqn-6.2} denotes the \ac{SINR} constraint for \me{\gamma_{l,k,n}}, and \eqref{eqn-6.3} gives an upper bound for the total interference seen by the user \me{k \in \mathcal{U}_b}, denoted by the variable \me{\beta_{l,k,n}}. The problem in \eqref{eqn-6} is known to be NP-hard even for the single antenna case \cite{wmmse_shi,kaleva2013primal}. The reformulation in \eqref{eqn-6} allows a tractable solution as presented below. First, we note that the constraints \eqref{eqn-4.3} are convex with involved variables. Thus, we only need to deal with \eqref{eqn-6.2} and \eqref{eqn-6.3}. Towards this end, we resort to the traditional coordinate descent technique by fixing the linear receivers, and find the optimal transmit beamformers. Recall the original coordinate descent method assumes that the optimization variables belong to disjoint sets (Cartesian product of sets, to be precise) \cite{xu2013block}.

By fixing the receivers, the problem now is to find optimal transmit beamformers for a given set of linear receivers which is a challenging task. We note that for fixed \me{\mvec{w}{l,k,n}}, \eqref{eqn-6.3} can be written as a \ac{SOC} constraint. Thus, the difficulty is due to the non-convexity in \eqref{eqn-6.2}. To arrive at a tractable formulation, we adopt the \ac{SCA} method to handle \eqref{eqn-6.2} by replacing the original non-convex constraint by the series of convex constraints. Note that the function \me{f(\tilde{\mbf{u}}_{l,k,n})} in \eqref{eqn-6.2} is convex for fixed \me{\mvec{w}{l,k,n}} since it is in fact the ratio between a quadratic form (of \me{\mvec{m}{l,k,n}}) over an affine function (of \me{\beta_{l,k,n}}) \cite{boyd2004convex}. According to the \ac{SCA} method, we relax \eqref{eqn-6.2} to a convex constraint in each iteration of the iterative procedure. Since \me{f(\tilde{\mbf{u}}_{l,k,n})} is convex, a concave approximation of \eqref{eqn-6.2} can be easily found by considering the first order approximation of \me{f(\tilde{\mbf{u}}_{l,k,n})} around the current operation point. For this purpose, let the real and imaginary component of the complex number \me{\mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n}} be represented by
\begin{subeqnarray} \label{eqn-wsrm-expr}
p_{l,k,n} &\triangleq& \Re \set{{\mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n}}} \\
q_{l,k,n} &\triangleq& \Im \set{{\mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n}}}
\end{subeqnarray}
and hence \me{f(\tilde{\mbf{u}}_{l,k,n})=(p_{l,k,n}^2 + q_{l,k,n}^2)/\beta_{l,k,n}}. Note that \me{p_{l,k,n}} and \me{q_{l,k,n}} are just symbolic notation and not the newly introduced optimization variables. In CVX \cite{grant2008cvx}, for example,  we declare \me{p_{l,k,n}} and \me{q_{l,k,n}} with the `\emph{expression}' qualifier. Suppose that the current value of \me{p_{l,k,n}} and \me{q_{l,k,n}} at a specific iteration are \me{\tilde{p}_{l,k,n}} and \me{\tilde{q}_{l,k,n}}, respectively. Using the first order Taylor approximation around the local point \me{ [ \, \tilde{p}_{l,k,n},\tilde{q}_{l,k,n},\tilde{\beta}_{l,k,n} \, ]^T}, we can approximate \eqref{eqn-6.2} by the following linear inequality constraint
\begin{equation}
\label{eqn-8}
2 \dfrac{\tilde{p}_{l,k,n}}{\tilde{\beta}_{l,k,n}} \left ( p_{l,k,n} - \tilde{p}_{l,k,n} \right ) + 2 \dfrac{\tilde{q}_{l,k,n}}{\tilde{\beta}_{l,k,n}} \left ( q_{l,k,n} - \tilde{q}_{l,k,n} \right ) + \dfrac{\tilde{p}_{l,k,n}^2 + \tilde{q}^2_{l,k,n}}{\tilde{\beta}_{l,k,n}} \left (1 - \dfrac{\beta_{l,k,n} - \tilde{\beta}_{l,k,n}}{\tilde{\beta}_{l,k,n}} \right ) \geq \gamma_{l,k,n}
\end{equation}
In summary, for the fixed linear receivers, the \ac{JSFRA} problem to find transmit beamformers is shown by
\begin{IEEEeqnarray}{rCl}\label{eqn-9}
\underset{\substack{\gamma_{l,k,n},p_{l,k,n},q_{l,k,n} \\ \mvec{m}{l,k,n}, \beta_{l,k,n}}}{\text{minimize}} &\quad & \| \tilde{\mbf{v}} \|_q \IEEEyessubnumber\label{eqn-9.1a} \\
\text{subject to} & \quad & \beta_{l,k,n} \geq  N_0\|\mvec{w}{l,k,n}\|^2 + \hspace{-0.75em} \sum_{(j,i) \neq (l,k)} \hspace{-0.75em} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \label{eqn-9.1c} \\
& \quad&\sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \text{tr} \, (\mvec{M}{k,n} \mvec{M}{k,n}^\herm) \leq P_{{\max}}, \fall b \IEEEyessubnumber \label{eqn-9.1d} \\
& \quad & \text{and } \eqref{eqn-8} \IEEEyessubnumber \label{eqn-9.1e}
\end{IEEEeqnarray}

Now, the optimal linear receivers for the fixed transmit precoders \me{\mbf{M}_{i,n} \, \forall i \in \mc{U}, \, \forall n \in \mc{C}} are obtained by minimizing \eqref{eqn-3} w.r.t \me{\mbf{w}_{l,k,n}} as
\begin{IEEEeqnarray}{rCl}\label{eqn-9--1}
\underset{\substack{\gamma_{l,k,n},p_{l,k,n},q_{l,k,n} \\ \mvec{w}{l,k,n}, \beta_{l,k,n}}}{\text{minimize}} &\quad & \| \tilde{\mbf{v}} \|_q \IEEEyessubnumber\label{eqn-9--1.1a} \\
\text{subject to} & \quad & \eqref{eqn-9.1c}, \eqref{eqn-9.1d}, \eqref{eqn-9.1e}, \text{ and } \eqref{eqn-8} \IEEEyessubnumber \label{eqn-9--1.1b}
\end{IEEEeqnarray}
Solving \eqref{eqn-9--1} using \ac{KKT} conditions, we obtain the following iterative expression for the receive beamformer \me{\mvec{w}{l,k,n}} as
\begin{subeqnarray} \label{opt-rx}
\mvec{R}{l,k,n} &=& \displaystyle \sum_{(j,i)\neq (l,k)} \mvec{H}{b_i,k,n} {\mbf{m}}_{j,i,n} {\mbf{m}}_{j,i,n}^\herm \mvec{H}{b_i,k,n}^\herm + N_0 \, \mathbf{I}_{N_R} \\
\mvec{w}{l,k,n}^{(i)} &=& \left ( \frac{\tilde{\beta}_{l,k,n} {\mbf{m}}_{l,k,n}^\herm \mvec{H}{b_k,k,n}^\herm \mbf{w}_{l,k,n}^{(i-1)} }{\|\mvec{w}{l,k,n}^{(i-1)} \mvec{H}{b_k,k,n} {\mbf{m}}_{l,k,n} \|^2} \right ) \, \mvec{R}{l,k,n}^{-1} \, \mvec{H}{b_k,k,n} {\mbf{m}}_{l,k,n}
\end{subeqnarray}
where \me{\mvec{w}{l,k,n}^{(i-1)}} is the receive beamformer from the earlier iteration, upon which the linear relaxation is performed for the nonconvex constraint in \eqref{eqn-9--1}. It can be seen that the optimal receiver expression in \eqref{opt-rx} is the scaled version of the \ac{MMSE} receiver, which is given by
\begin{IEEEeqnarray}{rCl}\label{eqn-10}
\mvec{w}{l,k,n} &=& \mathbf{R}^{-1}_{l,k,n} \; \mvec{H}{b_k,k,n} \; \mvec{m}{l,k,n} \IEEEyessubnumber \\
\mvec{R}{l,k,n} &=& \displaystyle \sum_{i\in \mc{U}} \sum_{j=1}^L \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} \mvec{m}{j,i,n}^\herm \mvec{H}{b_i,k,n}^\herm + N_0 \, \mathbf{I}_{N_R} \IEEEyessubnumber
\end{IEEEeqnarray}

>>>>

In this approach, we present an algorithm to solve \eqref{eqn-3} for the transmit precoders in a centralized way by using the idea of alternating optimization and successive convex approximation. Using \eqref{eq:SINR}, we can reformulate the problem defined in \eqref{eqn-3} as
\begin{IEEEeqnarray}{RCL}\label{eqn-6}
\underset{\substack{\gamma_{l,k,n},\mvec{m}{l,k,n},\\\beta_{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} & \quad & \|  \tilde{\mbf{v}}  \|_q \label{eqn-obj} \IEEEyessubnumber \\
\text{subject to}& \quad &\gamma_{l,k,n} \leq \dfrac{\left | \mvec{w}{l,k,n}^\herm \mvec{H}{l,k,n} \mvec{m}{l,k,n} \right |^2}{\beta_{l,k,n}} \triangleq f(\tilde{\mbf{u}}_{l,k,n}) \IEEEyessubnumber \label{eqn-6.2} \\
  & \quad & \beta_{l,k,n} \geq  N_0 \|\mvec{w}{l,k,n}\|^2 + \hspace{-0.75em} \sum_{(j,i) \neq (l,k)} \hspace{-0.75em} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \label{eqn-6.3} \\
  & \quad &\eqref{eqn-4.3} \IEEEyessubnumber
\end{IEEEeqnarray}
where \me{\tilde{\mbf{u}}_{l,k,n} \triangleq \{\mvec{w}{l,k,n}^\herm, \mvec{H}{b_k,k,n}, \mvec{m}{l,k,n},\beta_{l,k,n}\}} be the vector which needs to be identified for the optimal allocation. In this formulation, we relaxed the equality constraint in \eqref{eq:SINR} by the inequalities in \eqref{eqn-6.2} and \eqref{eqn-6.3}. However, this step is without loss of optimality leads to the same solution, since the inequalities in \eqref{eqn-6.2} and \eqref{eqn-6.3} are active for an optimal solution, following the same arguments as those in \cite{tran2012fast}. Intuitively, \eqref{eqn-6.2} denotes the \ac{SINR} constraint for \me{\gamma_{l,k,n}}, and \eqref{eqn-6.3} gives an upper bound for the total interference seen by the user \me{k \in \mathcal{U}_b}, denoted by the variable \me{\beta_{l,k,n}}. Similar to the \ac{WSRM} problem in \cite{tran2012fast}, the problem can be shown to be NP-hard even for the single antenna case. The reformulation in \eqref{eqn-6} allows a tractable solution as presented below. First, we note that the constraints \eqref{eqn-4.3} are convex with involved variables. Thus, we only need to deal with \eqref{eqn-6.2} and \eqref{eqn-6.3}. Towards this end, we resort to the traditional coordinate descent technique by fixing the linear receivers, and find the optimal transmit beamformers. Recall the original coordinate descent method assumes that the optimization variables belong to disjoint sets (Cartesian product of sets, to be precise) \cite{xu2013block}.

By fixing the receivers, the problem now is to find optimal transmit beamformers for a given set of linear receivers which is still a challenging task. We note that for fixed \me{\mvec{w}{l,k,n}}, \eqref{eqn-6.3} can be written as a \ac{SOC} constraint. Thus, the difficulty is due to the non-convexity in \eqref{eqn-6.2}. To arrive at a tractable formulation, we adopt the \ac{SCA} method to handle \eqref{eqn-6.2} by replacing the original non-convex constraint by the series of convex constraints. Note that the function \me{f(\tilde{\mbf{u}}_{l,k,n})} in \eqref{eqn-6.2} is convex for fixed \me{\mvec{w}{l,k,n}} since it is in fact the ratio between a quadratic form (of \me{\mvec{m}{l,k,n}}) over an affine function (of \me{\beta_{l,k,n}}) \cite{boyd2004convex}. According to the \ac{SCA} method, we relax \eqref{eqn-6.2} to a convex constraint in each iteration of the iterative procedure. Since \me{f(\tilde{\mbf{u}}_{l,k,n})} is convex, a concave approximation of \eqref{eqn-6.2} can be easily found by considering the first order approximation of \me{f(\tilde{\mbf{u}}_{l,k,n})} around the current operation point. For this purpose, let the real and imaginary component of the complex number \me{\mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n}} be represented by
\begin{subeqnarray} \label{eqn-wsrm-expr}
p_{l,k,n} &\triangleq& \Re \set{{\mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n}}} \\
q_{l,k,n} &\triangleq& \Im \set{{\mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n}}}
\end{subeqnarray}
and hence \me{f(\tilde{\mbf{u}}_{l,k,n})=(p_{l,k,n}^2 + q_{l,k,n}^2)/\beta_{l,k,n}}. Note that \me{p_{l,k,n}} and \me{q_{l,k,n}} are just symbolic notation and not the newly introduced optimization variables. In CVX \cite{grant2008cvx}, for example,  we declare \me{p_{l,k,n}} and \me{q_{l,k,n}} with the `\emph{expression}' qualifier. Suppose that the current value of \me{p_{l,k,n}} and \me{q_{l,k,n}} at a specific iteration are \me{\tilde{p}_{l,k,n}} and \me{\tilde{q}_{l,k,n}}, respectively. Using the first order Taylor approximation around the local point \me{ [ \, \tilde{p}_{l,k,n},\tilde{q}_{l,k,n},\tilde{\beta}_{l,k,n} \, ]^T}, we can approximate \eqref{eqn-6.2} by the following linear inequality constraint
\begin{equation}
\label{eqn-8}
2 \dfrac{\tilde{p}_{l,k,n}}{\tilde{\beta}_{l,k,n}} \left ( p_{l,k,n} - \tilde{p}_{l,k,n} \right ) + 2 \dfrac{\tilde{q}_{l,k,n}}{\tilde{\beta}_{l,k,n}} \left ( q_{l,k,n} - \tilde{q}_{l,k,n} \right ) + \dfrac{\tilde{p}_{l,k,n}^2 + \tilde{q}^2_{l,k,n}}{\tilde{\beta}_{l,k,n}} \left (1 - \dfrac{\beta_{l,k,n} - \tilde{\beta}_{l,k,n}}{\tilde{\beta}_{l,k,n}} \right ) \geq \gamma_{l,k,n}
\end{equation}
In summary, for the fixed linear receivers, the \ac{JSFRA} problem to find transmit beamformers is shown by
\begin{IEEEeqnarray}{rCl}\label{eqn-9}
\underset{\substack{\gamma_{l,k,n},p_{l,k,n},q_{l,k,n} \\ \mvec{m}{l,k,n}, \beta_{l,k,n}}}{\text{minimize}} &\quad & \| \tilde{\mbf{v}} \|_q \IEEEyessubnumber\label{eqn-9.1a} \\
\text{subject to} & \quad & \beta_{l,k,n} \geq  N_0\|\mvec{w}{l,k,n}\|^2 + \hspace{-0.75em} \sum_{(j,i) \neq (l,k)} \hspace{-0.75em} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \label{eqn-9.1c} \\
& \quad&\sum_{n = 1}^N \sum_{k \in \mathcal{U}_b}\sum_{l = 1}^L \text{tr} \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}}, \fall b \IEEEyessubnumber \label{eqn-9.1d} \\
& \quad & \text{and } \eqref{eqn-8} \IEEEyessubnumber \label{eqn-9.1e}
\end{IEEEeqnarray}

Now, the optimal linear receivers for the fixed transmit precoders \me{\mbf{m}_{j,i,n} \, \forall i \in \mc{U}, \, \forall n \in \mc{C}} are obtained by minimizing \eqref{eqn-3} w.r.t \me{\mbf{w}_{l,k,n}} as
\begin{IEEEeqnarray}{rCl}\label{eqn-9--1}
\underset{\substack{\gamma_{l,k,n},p_{l,k,n},q_{l,k,n} \\ \mvec{w}{l,k,n}, \beta_{l,k,n}}}{\text{minimize}} &\quad & \| \tilde{\mbf{v}} \|_q \IEEEyessubnumber\label{eqn-9--1.1a} \\
\text{subject to} & \quad & \eqref{eqn-9.1c}, \eqref{eqn-9.1d}, \eqref{eqn-9.1e}, \text{ and } \eqref{eqn-8} \IEEEyessubnumber \label{eqn-9--1.1b}
\end{IEEEeqnarray}
Solving \eqref{eqn-9--1} using \ac{KKT} conditions, we obtain the following iterative expression for the receive beamformer \me{\mvec{w}{l,k,n}} as
\begin{subeqnarray} \label{opt-rx}
\mvec{\widetilde{R}}{l,k,n} &=& \displaystyle \sum_{(j,i)\neq (l,k)} \mvec{H}{b_i,k,n} {\mbf{m}}_{j,i,n} {\mbf{m}}_{j,i,n}^\herm \mvec{H}{b_i,k,n}^\herm + N_0 \, \mathbf{I}_{N_R} \\
\mvec{w}{l,k,n}^{(i)} &=& \left ( \frac{\tilde{\beta}_{l,k,n} {\mbf{m}}_{l,k,n}^\herm \mvec{H}{b_k,k,n}^\herm \mbf{w}_{l,k,n}^{(i-1)} }{\|\mvec{w}{l,k,n}^{(i-1)} \mvec{H}{b_k,k,n} {\mbf{m}}_{l,k,n} \|^2} \right ) \, \mvec{\widetilde{R}}{l,k,n}^{-1} \, \mvec{H}{b_k,k,n} {\mbf{m}}_{l,k,n}
\end{subeqnarray}
where \me{\mvec{w}{l,k,n}^{(i-1)}} is the receive beamformer from the earlier iteration, upon which the linear relaxation is performed for the nonconvex constraint in \eqref{eqn-9--1}. It can be seen that the optimal receiver expression in \eqref{opt-rx} is the scaled version of the \ac{MMSE} receiver, which is given by
\begin{IEEEeqnarray}{rCl}\label{eqn-10}
\mvec{R}{l,k,n} &=& \displaystyle \sum_{i\in \mc{U}} \sum_{j=1}^L \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} \mvec{m}{j,i,n}^\herm \mvec{H}{b_i,k,n}^\herm + N_0 \, \mathbf{I}_{N_R} \IEEEyessubnumber \\
\mvec{w}{l,k,n} &=& \mathbf{R}^{-1}_{l,k,n} \; \mvec{H}{b_k,k,n} \; \mvec{m}{l,k,n} \IEEEyessubnumber
\end{IEEEeqnarray}
